Bayesian Exploration in Deep Reinforcement Learning

### 2022-08-27 papers/CDRL/sample_paper.tex
TEMPORARY DEPLOY COMMIT

Starting with a random variable \(S\) with a standard normal distribution
\(\N(0, I_{\lvert \A \rvert})\), we want a function \(v_\theta = v_{\theta_K}
\circ \dots \circ v_{\theta_2} \circ v_{\theta_1} \), such that
\(Z_\theta(\rvx, a) = v_\theta(S; \rvx)_a\). To learn \(v_\theta\) we note that
we can take the inverse transformation \(v^{-1}_\theta\) of the value
distribution to recover a random variable with the standard normal distribution,
i.e. \(S = v^{-1}_\theta(Z_\theta; \rvx)\). This further means that

\begin{equation}
  S^\prime \eqdist v^{-1}_\theta \left(\T v_\theta(S; \rvx); \rvx^\prime \right).
\end{equation}

We can use this

\begin{algorithm}
  \caption{Flow-RL Loss}\label{alg:LTM}
  \textbf{Require:} \(N, v, v^{-1}, \theta, \tilde{\theta}\)\\
  \textbf{input} \(\rvx, a, r, \rvx^\prime, \gamma\)\\
  \hspace*{1em} {\color{gray} \# Take \(N\) samples from the value distribution.}\\
  \hspace*{1em} \textbf{for} \(i = 1 \dots N\)\\
  \hspace*{2em} \(\epsilon_i \sim \N \left(0, I_{\lvert \A \rvert} \right)\)\\
  \hspace*{2em} \textbf{for} \(a^\prime \in \A\)\\
  \hspace*{3em} \(Z_{\tilde{\theta}}(\rvx^\prime, a^\prime)_i \gets v_{\tilde{\theta}}(\epsilon_i; \rvx^\prime)_{a^\prime}\)\\
  \hspace*{2em} \textbf{end}\\
  \hspace*{1em} \textbf{end}\\
  \\\hspace*{1em} {\color{gray} \# Compute greedy next action.}\\
  \hspace*{1em} \(a^* \gets \argmax_{a^\prime} \frac{1}{N} \sum_{i=1}^N Z_{\tilde{\theta}}(\rvx^\prime, a^\prime)_i\)\\
  \\\hspace*{1em} {\color{gray} \# Calculate loss terms for each sample.}\\
  \hspace*{1em} \textbf{for} \(i = 1 \dots N\)\\
  \hspace*{2em} {\color{gray} \# Apply the distributional Bellman operator.}\\
  \hspace*{2em} \(\hat{Z}_{\tilde{\theta}}(\rvx, a)_i \gets r + \gamma Z_{\tilde{\theta}}(\rvx^\prime, a^*)_i\)\\
  \hspace*{2em} {\color{gray} \# Calculate log-probability under \(\N(0,1)\).}\\
  \hspace*{2em} \(\delta_i \gets \frac{1}{2} v_{\theta}^{-1} 
  \left(\hat{Z}_{\tilde{\theta}}(\rvx, a)_i; \rvx \right)^2\)\\
  \hspace*{2em} {\color{gray} \# Calculate log-derivative of \(v^{-1}_\theta\).}\\
  \hspace*{2em} \(\rho_i \gets \log \left\lvert \frac{\partial v_{\theta}^{-1} \left(\hat{Z}_{\tilde{\theta}}(\rvx, a)_i; \rvx \right)}{\hat{Z}_{\tilde{\theta}}(\rvx, a)_i} \right\rvert\)\\
  \hspace*{1em} \textbf{end}\\
  \textbf{output} \(\frac{1}{N}\sum_{i=1}^N \delta_i - \rho_i\)
\end{algorithm}

>>>>

Starting with a random variable \(S\) with a standard normal distribution
\(\N(0, I_{\lvert \A \rvert})\), we want a function \(v_\theta = v_{\theta_K}
\circ \dots \circ v_{\theta_2} \circ v_{\theta_1} \), such that
\(Z_\theta(\rvx, a) = v_\theta(S; \rvx^\prime)_a\). To learn \(v_\theta\) we note that
we can take the inverse transformation \(v^{-1}_\theta\) of the value
distribution to recover a random variable with the standard normal distribution,
i.e. \(S = v^{-1}_\theta(Z_\theta; \rvx)\). This further means that

\begin{equation}
  S^\prime \eqdist v^{-1}_\theta \left(\T v_\theta(S; \rvx^\prime); \rvx \right).
\end{equation}

Because the inverse transformation of the value distribution given its state is
standard normal, we can learn the parameters \(\theta\) using the
change-of-variables formula from Equation~\ref{eq:change_var_form1}.
Algorithm~\ref{alg:LTM} shows how to calculate the loss for this scheme.

\begin{algorithm}
  \caption{Flow-RL Loss}\label{alg:LTM}
  \textbf{Require:} \(N, v, v^{-1}, \theta, \tilde{\theta}\)\\
  \textbf{input} \(\rvx, a, r, \rvx^\prime, \gamma\)\\
  \hspace*{1em} {\color{gray} \# Take \(N\) samples from the value distribution.}\\
  \hspace*{1em} \textbf{for} \(i = 1 \dots N\)\\
  \hspace*{2em} \(\epsilon_i \sim \N \left(0, I_{\lvert \A \rvert} \right)\)\\
  \hspace*{2em} \textbf{for} \(a^\prime \in \A\)\\
  \hspace*{3em} \(Z_{\tilde{\theta}}(\rvx^\prime, a^\prime)_i \gets v_{\tilde{\theta}}(\epsilon_i; \rvx^\prime)_{a^\prime}\)\\
  \hspace*{2em} \textbf{end}\\
  \hspace*{1em} \textbf{end}\\
  \\\hspace*{1em} {\color{gray} \# Compute greedy next action.}\\
  \hspace*{1em} \(a^* \gets \argmax_{a^\prime} \frac{1}{N} \sum_{i=1}^N Z_{\tilde{\theta}}(\rvx^\prime, a^\prime)_i\)\\
  \\\hspace*{1em} {\color{gray} \# Calculate loss terms for each sample.}\\
  \hspace*{1em} \textbf{for} \(i = 1 \dots N\)\\
  \hspace*{2em} {\color{gray} \# Apply the distributional Bellman operator.}\\
  \hspace*{2em} \(\hat{Z}_{\tilde{\theta}}(\rvx, a)_i \gets r + \gamma Z_{\tilde{\theta}}(\rvx^\prime, a^*)_i\)\\
  \hspace*{2em} {\color{gray} \# Calculate log-likelihood under \(\N(0,1)\).}\\
  \hspace*{2em} \(\delta_i \gets \frac{1}{2} v_{\theta}^{-1} 
  \left(\hat{Z}_{\tilde{\theta}}(\rvx, a)_i; \rvx \right)^2\)\\
  \hspace*{2em} {\color{gray} \# Calculate log-derivative of \(v^{-1}_\theta\).}\\
  \hspace*{2em} \(\rho_i \gets \log \left\lvert \frac{\partial v_{\theta}^{-1} \left(\hat{Z}_{\tilde{\theta}}(\rvx, a)_i; \rvx \right)}{\hat{Z}_{\tilde{\theta}}(\rvx, a)_i} \right\rvert\)\\
  \hspace*{1em} \textbf{end}\\
  \textbf{output} \(\frac{1}{N}\sum_{i=1}^N \delta_i - \rho_i\)
\end{algorithm}
